---
title: "YRWA Mapping - Fig. 1"
author: "Jessie Williamson"
date: "Revised 7/25/17"
output: html_document
---

######

One of several .rmd files associated with Williamson et al. 2019 "Community composition of sky-island parasites in Audubon's Warblers", International Journal for Parasitology: 10.1016/j.ijpara.2018.11.012. 

Script includes code for producing map Figure 1 in Williamson et al. 2019. 

Code originally written as a .Rmd in 2017-2019 but cleaned up a bit for GitHub upload on 10/31/20. 

#####

---

```{R, echo=FALSE}
# Set GLOBAL R chunk options here
#   (to hide this message add "echo=FALSE" to the code chunk options)

knitr::opts_chunk$set(comment = NA, message = FALSE, warning = FALSE, width = 100)
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 6)

#knitr::opts_chunk$set(cache = TRUE, autodep=TRUE)
knitr::opts_chunk$set(cache = TRUE, autodep=TRUE)
```

---

# Set WD
```{R}
rm(list=ls(all=TRUE)) # clear workspace 
setwd("/Users/Jessie/Dropbox (MSBbirds)/Rdirectory/YRWA_Malaria")
```

Mapping Resources:
http://www.gis-blog.com/r-raster-data-acquisition/ # This is most helpful tutorial
SRTM data search: http://srtm.csi.cgiar.org/SELECTION/inputCoord.asp
https://pakillo.github.io/R-GIS-tutorial/
http://www.molecularecologist.com/2012/09/making-maps-with-r/
http://www.spatialecology.com/beyer/assets/Beyer_Introduction_to_geospatial_analysis_in_R.pdf

NOTE: This code will run pretty slowly because these are clunky files. 


# Load in packages:
```{R}
library(maps) # contains functions I'll need
library(mapdata) # contains basic world map data 
library(maptools) # for shapefiles 
library(scales)
library(sp)  # classes for spatial data
library(raster)  # grids, rasters
library(rasterVis)  # raster visualisation
library(rgeos)
library(dismo)
library(ggmap)
```


# Load in the data:
```{R}
sites <- read.csv("YRWA_site_coordinates_mapping.csv") # Load in CSV w/ mean lat/lon for sites
# one column contains lat, one lon w/ coordinates in decimal degrees
```


Make a map of mean site coordinates:  
- Enter min and max lat/lon extent into CGIAR-CSI to get full map extent: http://srtm.csi.cgiar.org/SELECTION/inputCoord.asp
- Use this info to pull appropriate SRTM tiles 
```{r}
# download in chunks, later knit w/ mosaic. Adjust these to get AZ.
# srtm is dataset; returns srtm 90 m data
# lon pulls longitude of the srtm tile
# lat pulls latitude of srtm tile

srtm <- getData('SRTM',lon=-110,lat=33)
srtm2 <- getData('SRTM',lon=-110,lat=33)
srtm3 <- getData('SRTM',lon=-105,lat=33)
srtm4 <- getData('SRTM',lon=-110,lat=36)
srtm5 <- getData('SRTM',lon=-110,lat=36)
srtm6 <- getData('SRTM',lon=-105,lat=36)
srtmmosaic <- mosaic(srtm,srtm2,srtm3,srtm4,srtm5,srtm6,fun=mean)
usa <- getData('GADM',country='USA',level=1) # level 0 = blank; level 1 = county/state lines
# plot(srtmmosaic) # this just takes extra time 

# Plot points in teal on the grayscale map
plot(srtmmosaic,xlim=c(-112,-101.7),ylim=c(31,37.5),col=rev(gray.colors(10,0,1))) # this zooms in to the area of NM 
plot(usa,add=TRUE) # adds boundary of NM 
points(sites$mean_lon,sites$mean_lat,pch=21,bg="#54D3BF",cex=2.4) # Plots points in teal; adjust "bg" for color 
# cex should be set to 2.5 for manuscript fig, and ~4 for graphical abstract

# BLM Teal color: #54D3BF
# Writer for GIS tutorial above says to drag R Studio windows to eliminate white space on either side of the plot
# although when I do that the dimensions become distorted - takes careful finagling

# add text labels for state names 
# text(-108.8,31.5,"ARIZONA",cex=0.7,pos=2,font=2)

### also try this but clip AZ: e.g. adjust your -115 to be less to see if this helps; maybe even remove these layers 
``` 


# old mapping code - didn't end up using? 
```{r}
srtm <- getData('SRTM',lon=-106,lat=36) 
plot(srtm)
srtm2 <- getData('SRTM',lon=-106,lat=33) 
plot(srtm2)
srtm3 <- getData('SRTM',lon=-102,lat=33)
srtm4 <- getData('SRTM',lon=-102,lat=36)
srtm5 <- getData('SRTM',lon=-108,lat=36)
srtm6 <- getData('SRTM',lon=-108,lat=33)
srtmmosaic <- mosaic(srtm,srtm2,srtm3,srtm4,srtm5,srtm6,fun=mean)
usa <- getData('GADM',country='USA',level=1) # level 0 = blank; level 1 = county/state lines

# Add points to map of US (full color)
plot(srtmmosaic,xlim=c(-110,-102),ylim=c(31.1,37.25),col=rev(gray.colors(10,0,1))) # Plots a map of NM & AZ; adjust these to clip map
plot(usa,add=TRUE)
points(sites$mean_lon,sites$mean_lat,pch=21,bg="darkgray",cex=1.4) # This plots sampling points on map

# Deliminations of raw plot w/out state boundaries, etc. 
plot(srtmmosaic,xlim=c(-112,-102),ylim=c(31,38))
```


----

# End 


